Testing and Isolation Efficacy: Insights from a Simple Epidemic Model

### 2021-04-01 SIR_manuscript.tex
proving that R0 decreases wrt rho without using Tylor appr

The basic reproduction number, $\Rnum$, was calculated by using the next generation matrix method developed by \cite{van2002reproduction}. $\Rnum$ is
\begin{equation}
\label{R0}
\Rnum= \frac{\beta}{\gamma} (1-\Delta), 
\end{equation}
where
\begin{equation}
\begin{aligned}
\label{eq:abc}
\Delta=& (A~ \theta\_w + B~ \theta\_c)~ C\\
A=& \gamma \big( (\gamma \omega+ \omega p_I F_I+\gamma^2) x_n^* + (\omega+\gamma) F_I \big)\\
B=& \omega p_I F_I (\gamma x_u^*+\omega) \\ 
C=& \frac{1}{(\omega+\gamma) \big(\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I) \big)}.
\end{aligned}
\end{equation}

@@

\begin{enumerate}
\item \label{p1:eta} increasing the relative probabilities of transmission for tested individuals, increases $\Rnum$;
$$\partial{\Rnum}/\partial{\theta\_c} \leq 0 \text{~and~} \partial{\Rnum}/\partial{\theta\_w} \leq 0.$$ 
\item \label{p1:rho} higher rate of testing reduces $\Rnum$;
$$\partial{\Rnum}/\partial{\rho} \leq 0 \text{~when~} \rho \approx 0.$$
\item \label{p1:omega} The rate of test return may result in the disease becomes extinct or not;
$$\partial{\Rnum}/\partial{\omega} \text{~can be positive or negative when~} \rho \approx 0.$$
\end{enumerate}

@@

% %%%%%%%
\subsection{Taylor Approximation of $\Rnum$ at $\rho=0$ }
The basic reproduction number, $\Rnum$, close to $\rho=0$ can be approximated linearly in $\rho$ by using Taylor approximation. it follows
\begin{equation}
\label{eq:R0appr}
\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(-\theta\_w)(\gamma w_S+\omega w_I) + (-\theta\_c)p_I w_I \omega^2 \Big) + \mathcal{O}(\rho^2).
\end{equation}

Also, to analyze the influence of $\omega$ on $\Rnum$, the approximation \eqref{eq:R0appr} was used. Then it is straight forward to have
\begin{equation}
\label{Rom}
\partial{\Rnum}/\partial{\omega}=  \frac{-\beta \rho}{\gamma w_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),
\end{equation}
where $a=(-\theta\_w)w_I-(-\theta\_c)p_I w_I = ((s-p_I)(1-\theta\_c) + (p_I-1)) w_I$, $b=2(-\theta\_w)\gamma w_S$ and $c=(-\theta\_w)\gamma^2 w_S$.
Given that $0 \leq \theta\_w \leq \theta\_c \leq 1 $, one can easily derive $b\leq 0$ and $c \leq 0$.

Note that in general, the necessary and sufficient condition for $a \geq 0$ is $(s-p_I) (1-\theta\_c) \geq (1-p_I)$, where $s=\frac{(1-\theta\_w)}{(1-\theta\_c)} \geq 1$.

In the case of ``perfect isolation'', i.e., when $\theta\_w=1$ and consequently $\theta\_c=1$, it is straightforward to see that $a \leq 0$, $b<0$ and $c<0$. Thus, $\partial{\Rnum}/\partial{\omega} \geq 0$. 
As an example, in the case of a very accurate testing regime,  i.e., $P_I=1$, $a \geq 0$ is achieved. If $a\geq 0$, the quadratic expression in \eqref{Rom}, has Real roots. Assuming that $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{\Rnum}/\partial{\omega}$. Thus, $\partial{\Rnum}/\partial{\omega}>0$ for $0<\omega<\omega_2$ and  $\partial{\Rnum}/\partial{\omega}<0$ for $\omega>\omega_2$.

>>>>

The basic reproduction number, $\Rnum$, was calculated by using the next generation matrix method developed by \cite{van2002reproduction}. $\Rnum$ is
\begin{equation}
\label{R0}
\Rnum= \frac{\beta}{\gamma} (1-\Delta), 
\end{equation}
where
\begin{equation}
\begin{aligned}
\label{eq:abc}
\Delta=& (A~ \theta\_w + B~ \theta\_c) / C\\
A=& \gamma \Big( \big(\gamma(\omega+\gamma) + \omega p_I F_I \big) x_n^* + (\omega+\gamma) F_I \Big)\\
B=& \omega p_I F_I (\gamma x_u^*+\omega) \\ 
C=& (\omega+\gamma) \Big(\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I) \Big).
\end{aligned}
\end{equation}

@@

\begin{enumerate}
\item \label{p1:eta} increasing the relative probabilities of transmission for tested individuals, increases $\Rnum$;
$$\partial{\Rnum}/\partial{\theta\_c} \leq 0 \text{~and~} \partial{\Rnum}/\partial{\theta\_w} \leq 0.$$ 
\item \label{p1:rho} higher rate of testing reduces $\Rnum$;
$$\partial{\Rnum}/\partial{\rho} \leq 0.$$
\item \label{p1:omega} The rate of test return may result in the disease becomes extinct or not;
$$\partial{\Rnum}/\partial{\omega} \text{~can be positive or negative}.$$
\end{enumerate}

@@

% %%%%%%%
\subsection{More testing rate reduces $\Rnum$}

 The expression of the basic reproduction number $\Rnum$, given in Eqs. \eqref{R0}, makes it straight  forward to conclude $\frac{\partial \Delta}{\partial \rho} > 0$ and thus $\frac{\partial \Rnum}{\partial \rho} < 0$. 
Given the fractional form of $\Delta$, it is easy to show that $\partial A/\partial \rho >0$, $\partial B/\partial \rho >0$ and $\partial C/\partial \rho <0$. Note that we use the fact that
$\partial F_I/\partial \rho >0$, $\partial S\_n^*/\partial \rho >0$ and $\partial S\_u^*/\partial \rho <0$.
